First steps in synthetic algebraic geometry

### 2017-03-17 snippets/qcoh.tex
More details on the counterexample

\begin{thm}\label{thm:qcoh-big-char}
Let~$E \in \Zar(S)$ be an~$\affl$-module.
If~$E$ is quasicoherent, that is of
the form~$\E_0^\Zar$ for some quasicoherent~$\O_S$-module~$\E_0$,
then~$E$ is synthetically quasicoherent from the internal point of view of~$\Zar(S)$.
The converse holds in any of the following situations:
\begin{enumerate}
\item The site used to construct~$\Zar(S)$ is one of the parsimonious sites.
\item The base scheme~$S$ is concentrated (quasicompact and quasiseparated) and
the functor~$E$ maps directed limits of inverse systems of~$S$-schemes with
affine transition morphisms to colimits in~$\Set$.
\end{enumerate}
\end{thm}

@@

By Proposition~\ref{XXX}, the internal expression~$\Spec(A)$ is the functor of
points of~$\RelSpek_S \A_0$. For any~$S$-scheme~$f : T \to S$ contained in the site
used to define~$\Zar(S)$, we consider the fiber product
\[ \xymatrix{
  \RelSpek_T(f^*\A_0) \ar[r]^{f'} \ar[d]_{p'} & \RelSpek_S\A_0 \ar[d]^p \\
  T \ar[r]_f & S.
} \]
Since~$\RelSpek_T(f^*\A_0) \to S$ is contained in the site (for any of our
admissible sites), we may conclude using the following chain of isomorphisms:
\begin{align*}
  [\Spec(A), E](T) &\cong
  \Hom_{\Zar(S)}(\ul{T}, [\Spec(A), E])
  \cong \Hom_{\Zar(S)}(\ul{T} \times \Spec(A), E) \\
  &\cong \Hom_{\Zar(S)}(\underline{T \times_S \RelSpek_S\A_0}, E)
  \cong E(\RelSpek_T(f^*\A_0)) \\
  &\cong \Gamma(\RelSpek_T(f^*\A_0), (p')^* f^* \E_0)
  \cong \Gamma(T, (p')_* (p')^* f_* \E_0) \\
  &\cong \Gamma(T, f^*\E_0 \otimes_{\O_T} f^*\A_0)
  \cong \Gamma(T, (\E_0 \otimes_{\O_S} \A_0)^\Zar) \\
  &\cong \Gamma(T, \E_0^\Zar \otimes_\affl \A_0^\Zar)
  \cong \Gamma(T, E \otimes_\affl A).
\end{align*}
The antepenultimate isomorphism is because pullback of modules in~$\Sh(S)$ to
modules in~$\Sh(T)$ commutes with tensor product. The penultimate isomorphism
is because pullback of a sheaf in~$\Sh(S)$ to a sheaf in~$\Zar(S)$ commutes
with tensor product (Lemma~\ref{lemma:zar-tensor-product-commutes}).

@@

A natural candidate for an quasicoherent~$\O_S$-module~$\E_0$ with~$E \cong
\E_0^\Zar$ is~$\E_0 \defeq E|_{\Sh(S)}$. We'll show that this is indeed true.
Let~$f : T \to S$ be any~$S$-scheme contained in the site used to
define~$\Zar(S)$. We assume for the time being that~$f$ is of finite
presentation and affine, so~$T \cong \RelSpek_S \A_0$ for some finitely
presented~$\O_S$-algebra~$\A_0$. We want to verify that the canonical morphism
\begin{equation}\label{eqn:want-iso2}\tag{§}
  f^*(E|_{\Sh(S)}) \longrightarrow E|_{\Sh(T)}
\end{equation}
is an isomorphism. Since the functor~$f_*$ from quasicoherent~$\O_T$-modules to
quasicoherent~$\O_S$-modules is fully faithful (the morphism~$f$ being affine)
and domain and codomain of that morphism are quasicoherent, it suffices to
verify that its image under~$f_*$ is an isomorphism. This image is the
canonical morphism
\[ E|_{\Sh(S)} \otimes_{\O_S} \A_0 \longrightarrow f_*(E|_{\Sh(T)}). \]
The assumption of synthetic quasicoherence, applied to the~$\affl$-algebra~$A
\defeq \A_0^\Zar$, shows that this morphism is an isomorphism.

@@

\begin{rem}The condition in Scholium~\ref{scholium} that~$\Spec(A)$ is
representable by an object of the site used to define~$\Zar(S)$ cannot be
dropped. For instance, if we employ the parsimonious sites and consider~$S =
\Spec \ZZ$,~$A = \K_S^\Zar$, and~$E = \affl$, then~$\Spec(A) = \emptyset$.
Therefore the codomain of the canonical morphism is the zero algebra, but the
domain is not.
\end{rem}

>>>>

\begin{thm}\label{thm:qcoh-big-char}
Let~$E \in \Zar(S)$ be an~$\affl$-module.
If~$E$ is quasicoherent, that is of
the form~$(\E_0)^\Zar$ for some quasicoherent~$\O_S$-module~$\E_0$,
then~$E$ is synthetically quasicoherent from the internal point of view of~$\Zar(S)$.
The converse holds in any of the following situations:
\begin{enumerate}
\item The site used to construct~$\Zar(S)$ is one of the parsimonious sites.
\item The base scheme~$S$ is concentrated (quasicompact and quasiseparated) and
the functor~$E$ maps directed limits of inverse systems of~$S$-schemes with
affine transition morphisms to colimits in~$\Set$.
\end{enumerate}
\end{thm}

@@

By Proposition~\ref{XXX}, the internal expression~$\Spec(A)$ is the functor of
points of~$\RelSpek_S \A_0$. For any~$S$-scheme~$f : T \to S$ contained in the site
used to define~$\Zar(S)$, we consider the fiber product
\[ \xymatrix{
  \RelSpek_T(f^*\A_0) \ar[r]^{f'} \ar[d]_{p'} & \RelSpek_S\A_0 \ar[d]^p \\
  T \ar[r]_f & S.
} \]
Since~$\RelSpek_T(f^*\A_0) \to S$ is contained in the site (for any of our
admissible sites), we may conclude using the following chain of isomorphisms:
\begin{align*}
  [\Spec(A), E](T) &\cong
  \Hom_{\Zar(S)}(\ul{T}, [\Spec(A), E])
  \cong \Hom_{\Zar(S)}(\ul{T} \times \Spec(A), E) \\
  &\cong \Hom_{\Zar(S)}(\underline{T \times_S \RelSpek_S\A_0}, E)
  \cong E(\RelSpek_T(f^*\A_0)) \\
  &\cong \Gamma(\RelSpek_T(f^*\A_0), (p')^* f^* \E_0)
  \cong \Gamma(T, (p')_* (p')^* f_* \E_0) \\
  &\cong \Gamma(T, f^*\E_0 \otimes_{\O_T} f^*\A_0)
  \cong \Gamma(T, (\E_0 \otimes_{\O_S} \A_0)^\Zar) \\
  &\cong \Gamma(T, (\E_0)^\Zar \otimes_\affl (\A_0)^\Zar)
  \cong \Gamma(T, E \otimes_\affl A).
\end{align*}
The antepenultimate isomorphism is because pullback of modules in~$\Sh(S)$ to
modules in~$\Sh(T)$ commutes with tensor product. The penultimate isomorphism
is because pullback of a sheaf in~$\Sh(S)$ to a sheaf in~$\Zar(S)$ commutes
with tensor product (Lemma~\ref{lemma:zar-tensor-product-commutes}).

@@

A natural candidate for an quasicoherent~$\O_S$-module~$\E_0$ with~$E \cong
(\E_0)^\Zar$ is~$\E_0 \defeq E|_{\Sh(S)}$. We'll show that this is indeed true.
Let~$f : T \to S$ be any~$S$-scheme contained in the site used to
define~$\Zar(S)$. We assume for the time being that~$f$ is of finite
presentation and affine, so~$T \cong \RelSpek_S \A_0$ for some finitely
presented~$\O_S$-algebra~$\A_0$. We want to verify that the canonical morphism
\begin{equation}\label{eqn:want-iso2}\tag{§}
  f^*(E|_{\Sh(S)}) \longrightarrow E|_{\Sh(T)}
\end{equation}
is an isomorphism. Since the functor~$f_*$ from quasicoherent~$\O_T$-modules to
quasicoherent~$\O_S$-modules is fully faithful (the morphism~$f$ being affine)
and domain and codomain of that morphism are quasicoherent, it suffices to
verify that its image under~$f_*$ is an isomorphism. This image is the
canonical morphism
\[ E|_{\Sh(S)} \otimes_{\O_S} \A_0 \longrightarrow f_*(E|_{\Sh(T)}). \]
The assumption of synthetic quasicoherence, applied to the~$\affl$-algebra~$A
\defeq (\A_0)^\Zar$, shows that this morphism is an isomorphism.

@@

\begin{rem}The condition in Scholium~\ref{scholium} that~$\Spec(A)$ is
representable by an object of the site used to define~$\Zar(S)$ cannot be
dropped. For instance, if we employ the parsimonious sites and consider~$S =
\Spec \ZZ$,~$E = \affl$, and $A = \K_S^\Zar$ (where~$\K_S$ is the sheaf of
rational functions on~$S$, which in this case is the constant sheaf~$\ul{\QQ}$),
then~$\Spec(A)$ is the functor of points of the~$\ZZ$-scheme~$\Spec(\QQ)$.
This functor coincides with the functor of points of the empty~$\ZZ$-scheme on
the parsimonious sites; therefore~$\Spec(A) = \emptyset$ from the internal point
of view. Thus the codomain of the canonical morphism is the zero algebra, but the
domain is not.
\end{rem}
